
### load the data

load the data through `read.csv`
```{r}
data <- read.csv("activity.csv", stringsAsFactors = FALSE)
data$date <- as.Date(data$date, "%Y-%m-%d")
str(data)
```

### What is mean total number of steps taken per day?

use tidyverse to summarise and visualization the data

#### Calculate the total number of steps taken per day
```{r}
library(tidyverse)
(totalCount <- data %>% group_by(date) %>% summarise(totalCount = sum(steps, na.rm = T)))
ggplot(totalCount, aes(x =totalCount)) + geom_histogram(color = "black", fill = "white") + theme_bw()
```

#### Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean(totalCount$totalCount)
median(totalCount$totalCount)
```

### What is the average daily activity pattern?

#### Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
(dataMeanPerInterval <- data %>% group_by(interval) %>% summarise(mean = mean(steps, na.rm = T)))
plot(dataMeanPerInterval$interval, dataMeanPerInterval$mean, type = "l")
```

#### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
dataMeanPerInterval$interval[which.max(dataMeanPerInterval$mean)]
```

### Imputing missing values

#### Calculate and report the total number of missing values in the dataset 

```{r}
sum(is.na(data$steps))
```

#### filling in all of the missing values in the dataset.
using the mean for that 5-minute interval to fill the dataset

```{r}
newdata <- data
for(i in dataMeanPerInterval$interval){
  newdata$steps[is.na(newdata$steps) & newdata$interval == i] <- dataMeanPerInterval$mean[dataMeanPerInterval$interval == i]
}
```
#### Make a histogram of the total number of steps taken each day

```{r}
(totalCount_new <- newdata %>% group_by(date) %>% summarise(totalCount = sum(steps, na.rm = T)))
ggplot(totalCount_new, aes(x =totalCount)) + geom_histogram(color = "black", fill = "white") + theme_bw()
mean(totalCount_new$totalCount)
median(totalCount_new$totalCount)
```

these values differ from the estimates from the first part of the assignment. 

### Are there differences in activity patterns between weekdays and weekends?

```{r}
newdata$week <- ifelse(weekdays(newdata$date) %in% c("Saturday", "Sunday"), "weekends", "weekdays")
(dataMeanIntervWeekday  <- newdata %>% group_by(interval, week) %>% summarise(mean = mean(steps, na.rm = T)))
ggplot(dataMeanIntervWeekday, aes(x = interval, y = mean)) + geom_line() + facet_grid(week ~ .)
```

